---
format: 
  pdf:
    include-in-header:
      text: |
        \addtokomafont{disposition}{\rmfamily}
        \usepackage{fancyhdr}
        \usepackage{setspace}
        \doublespacing
---

\raggedright
\pagestyle{fancy}
\fancyhf{}
\fancyfoot[R]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0.382pt}
\begin{center} \includegraphics[width=0.3\columnwidth]{images} \end{center}
\begin{center}
  {\Large Consumer Sentiments VS. Economy Realities\\
  Longitudinal Analysis of Changing Consumer Perceptions in Relation to Employment\par}
\end{center}
\begin{center} {Zupeng Zeng \& Troy (Shengkun) Liu} \end{center}
\begin{center} {December 2025} \end{center}
\newpage

# Socio-Economic Background

| 

\newpage

\begin{center} {\large \textbf{Exploratory Data Analysis Executive Summary}\par} \end{center}

-   **Project Objective**
-   **Data Source**
-   **Data Reliability**
-   **THEME FOUND123**
-   **Limitations of the Analysis** 

\newpage

# Research Questions

-   

-   

# Data Source and Assumptions

# Data Cleaning Process

# Notable Findings

***Finding 1.*** Finding Theme

State the finding in a few lines (See Visualization 1).

***Finding 2.*** (See Visualization 2).

***Finding 3.*** (See Visualization 3).

\newpage

# Visualizations

Working repo could be fount at: https://github.com/zzeng05/ZENG1-LIU2-727FINAL-scaVSeply.git

```{r setup}
#| warning: false
#| echo: false
#| message: false
# Import Libraries
library(tidyverse)
library(lubridate)
library(rvest)
library(zoo)        # for rolling correlations
library(broom)      # for regression summaries
library(ggplot2)
library(stringr)
library(readxl)
library(httr)
library(jsonlite)
library(purrr)

```

```{r}
# Fetch SCA Data
cs_url <- "https://data.sca.isr.umich.edu/data-archive/mine.php#"

#function to fetch any sca table from data site
get_sca_table <- function(table_num,
                               from_year = 2008,
                               to_year   = 2025,
                               freq      = "monthly") {
  
  body_list <- list(
    table = as.character(table_num),
    format = "html",
    from   = as.character(from_year),
    to     = as.character(to_year),
    freq   = freq
  )
  
  res <- POST(cs_url, body = body_list, encode = "form")
  stop_for_status(res)
  
  page <- read_html(res)
  
  page %>%
    html_node("div.output table") %>%
    html_table(fill = TRUE)
}
```

```{r}
#fetch & clean table1: Consumer Sentiment Index
CS <- get_sca_table(1) %>%
  # same logic as your cs_tables[[2]] cleaning
  rename(
    month_chr = X1,
    year_chr  = X2,
    cs_chr    = X3
  ) %>%
  slice(-1) %>%  # drop the header row inside the table
  mutate(
    month = as.integer(month_chr),
    year  = as.integer(year_chr),
    cs    = as.numeric(cs_chr),
    date  = ymd(sprintf("%04d-%02d-01", year, month))
  ) %>%
  arrange(date) %>%
  select(date, cs, year, month)

head(CS)
```

```{r}
#fetch & clean table30: Expected Change in Unemployment During the Next Year
UNEPLY1 <- get_sca_table(30) %>%
  { setNames(., as.character(unlist(.[1, ]))) } %>%
  slice(-1) %>%
  mutate(
    Month = as.integer(Month),
    Year  = as.integer(Year)
  ) %>%
  mutate(
    across(
      -c(Month, Year),
      ~ suppressWarnings(as.numeric(.))
    ),
    date = as.Date(sprintf("%04d-%02d-01", Year, Month))
  ) %>%
  relocate(date)

head(UNEPLY1)
```

```{r}
UNEPLY_supp <- UNEPLY1 %>%
  transmute(
    date,
    less_unemp  = Less,
    same_unemp  = Same,
    more_unemp  = More,
    dk_unemp    = `DK; NA`,
    rel_unemp   = Relative,
    net_unemp_expect = less_unemp - more_unemp
  )

head(UNEPLY_supp)

UNEPLY_long <- UNEPLY_supp %>%
  select(date, less_unemp, same_unemp, more_unemp) %>%
  pivot_longer(
    cols = -date,
    names_to  = "expectation",
    values_to = "share"
  )

UNEPLY_long <- UNEPLY_long |>
  dplyr::mutate(
    expectation = factor(
      expectation,
      levels = c("less_unemp", "same_unemp", "more_unemp"),
      labels = c("Less unemployment", "Same", "More unemployment")
    )
  )

ggplot(UNEPLY_long, aes(x = date, y = share, fill = expectation)) +
  geom_area(alpha = 0.618) +
  scale_fill_manual(
    values = c(
      "Less unemployment" = "darkgreen",
      "Same"              = "grey70",
      "More unemployment" = "firebrick"
    ),
    guide = guide_legend(reverse = FALSE)
  ) +
  labs(
    title = "Expected change in unemployment over the next year",
    x     = "Date",
    y     = "Percent of respondents",
    fill  = NULL
  ) +
  theme_minimal() +
  theme(
    panel.grid = element_blank()
  )
```

```{r}
# Fetch BLS Data

bls_key <- "554489bce3f14059aaa2dbb976d62372" #need to change back to path before submission!!!

# BLS series IDs:
# - LNS14000000: Unemployment rate (CPS, seasonally adjusted)
# - CES0000000001: All employees, total nonfarm (CES, thousands, SA)
series_ids <- c("LNS14000000", "CES0000000001")

# Request monthly data from 2008 to latest
res <- POST(
  url   = "https://api.bls.gov/publicAPI/v2/timeseries/data/",
  body  = list(
  seriesid        = series_ids,
  startyear       = "2008",
  endyear         = "2025",
  registrationkey = bls_key),
  encode = "json"
)

stop_for_status(res)

bls_list <- fromJSON(content(res, as = "text", encoding = "UTF-8"), simplifyDataFrame = TRUE)

# Transform BLS data to long format
series_tbl <- bls_list$Results$series

bls_long <- map_dfr(seq_len(nrow(series_tbl)), function(i) {
  series_id <- series_tbl$seriesID[i]
  dat       <- series_tbl$data[[i]]   #data.frame for the series
  
  as_tibble(dat) %>%
    transmute(
      series_id = series_id,
      year      = as.integer(year),
      period,
      value     = as.numeric(value)
    ) %>%
    # keep monthly observations (M01–M12)
    filter(str_starts(period, "M")) %>%
    mutate(
      month = as.integer(str_remove(period, "M")),
      date  = ymd(sprintf("%04d-%02d-01", year, month))
    )
})

head(bls_long)
```

```{r}
# CPS unemployment rate
UNEPLOY_R <- bls_long %>%
  filter(series_id == "LNS14000000") %>%
  arrange(date) %>%
  transmute(
    date,
    unrate = value
  )

# CES total nonfarm employment & monthly job change
Job_V <- bls_long %>%
  filter(series_id == "CES0000000001") %>%
  arrange(date) %>%
  transmute(
    date,
    nonfarm_emp = value,
    job_change  = nonfarm_emp - lag(nonfarm_emp)
  )

head(UNEPLOY_R)
head(Job_V)
```

## Visualization 1. Lagged Time-Series of Consumer Sentiment and Unemployment Rate/Job Change

```{r}
#| echo: false
#| warning: false
# Merge SCA and BLS series into one panel
macro <- CS %>%
  select(date, cs) %>%
  left_join(UNEPLOY_R, by = "date") %>%
  left_join(Job_V,     by = "date") %>%
  arrange(date)

head(macro)
```

```{r}
#| warning: false
#| message: false
# Function to plot Consumer Sentiment vs Unemployment Rate with dual y-axes
cs_vs_unemp <- function(data,
                                  lag_months   = 3,
                                  cs_limits    = c(30, 120),   # left y-axis
                                  unemp_limits = c(0, 15)) {   # right y-axis
  
  df <- data %>%
    arrange(date) %>%
    mutate(unrate_lead = lead(unrate, lag_months)) %>%
    filter(!is.na(cs), !is.na(unrate_lead))
  
  # unpack limits
  cs_min   <- cs_limits[1]
  cs_max   <- cs_limits[2]
  ur_min   <- unemp_limits[1]
  ur_max   <- unemp_limits[2]
  
  cs_range <- cs_max - cs_min
  ur_range <- ur_max - ur_min
  
  # scale unemployment into CS scale for plotting
  df <- df %>%
    mutate(
      unrate_scaled = cs_min + (unrate_lead - ur_min) * (cs_range / ur_range)
    )
  
  # dynamic label for the legend
  unemp_label <- paste0("Unemp (t + ", lag_months, "m)")
  
  ggplot(df, aes(x = date)) +
    geom_line(aes(y = cs, color = "Consumer Sentiment")) +
    geom_line(aes(y = unrate_scaled, color = unemp_label)) +
    scale_y_continuous(
      name   = "Consumer Sentiment Index",
      limits = cs_limits,
      sec.axis = sec_axis(
        trans = ~ (.- cs_min) * (ur_range / cs_range) + ur_min,
        name  = paste0("Unemployment rate at t + ", lag_months, " months")
      )
    ) +
    scale_color_manual(
      values = setNames(
        c("darkblue", "firebrick"),
        c("Consumer Sentiment", unemp_label)
      )
    ) +
    labs(
      title = paste0("Consumer Sentiment vs Future Unemployment (", lag_months, "-month lead)"),
      x     = "Date",
      color = NULL
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom"
    )
}

cs_vs_unemp (macro, lag_months = 0)
cs_vs_unemp (macro, lag_months = 6)
cs_vs_unemp (macro, lag_months = 5)
cs_vs_unemp (macro, lag_months = 7)
cs_vs_unemp (macro, lag_months = 8)
```
```{r}
#| warning: false
#| message: false
cs_vs_unemp <- function(data,
                        lag_months   = 3,
                        cs_limits    = c(30, 120),   # left y-axis
                        unemp_limits = c(0, 15)) {   # right y-axis
  
  df <- data %>%
    arrange(date) %>%
    mutate(unrate_lead = lead(unrate, lag_months)) %>%
    filter(!is.na(cs), !is.na(unrate_lead))
  
  # unpack limits
  cs_min   <- cs_limits[1]
  cs_max   <- cs_limits[2]
  ur_min   <- unemp_limits[1]
  ur_max   <- unemp_limits[2]
  
  cs_range <- cs_max - cs_min
  ur_range <- ur_max - ur_min
  
  # scale unemployment into CS scale for plotting
  df <- df %>%
    mutate(
      unrate_scaled = cs_min + (unrate_lead - ur_min) * (cs_range / ur_range)
    )
  
  # dynamic label for the legend
  unemp_label <- paste0("Unemp (t + ", lag_months, "m)")
  
  ggplot(df, aes(x = date)) +
    # raw lines (lighter)
    geom_line(aes(y = cs, color = "Consumer Sentiment"), alpha = 0.4) +
    geom_line(aes(y = unrate_scaled, color = unemp_label), alpha = 0.4) +
    
    # smoothed lines (highlighted)
    geom_smooth(aes(y = cs, color = "Consumer Sentiment"),
                se = FALSE, span = 0.2, linewidth = 0.9) +
    geom_smooth(aes(y = unrate_scaled, color = unemp_label),
                se = FALSE, span = 0.2, linewidth = 0.9) +
    
    scale_y_continuous(
      name   = "Consumer Sentiment Index",
      limits = cs_limits,
      sec.axis = sec_axis(
        trans = ~ (.- cs_min) * (ur_range / cs_range) + ur_min,
        name  = paste0("Unemployment rate at t + ", lag_months, " months")
      )
    ) +
    scale_color_manual(
      values = setNames(
        c("darkblue", "firebrick"),
        c("Consumer Sentiment", unemp_label)
      )
    ) +
    labs(
      title = paste0("Consumer Sentiment vs Future Unemployment (", lag_months, "-month lead)"),
      x     = "Date",
      color = NULL
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom"
    )
}

cs_vs_unemp (macro, lag_months = 0)
cs_vs_unemp (macro, lag_months = 6)
cs_vs_unemp (macro, lag_months = 5)
cs_vs_unemp (macro, lag_months = 7)
cs_vs_unemp (macro, lag_months = 8)
```



```{r}
#| warning: false
#| message: false
# function to plot Consumer Sentiment vs Job Change with dual y-axes
cs_vs_jobs <- function(data,
                                 lag_months  = 3,
                                 cs_limits   = c(40, 120), # left y-axis
                                 jobs_limits = c(-1500, 1500)) {# right y-axis (thousands)
  
  df <- data %>%
    arrange(date) %>%
    mutate(jobchg_lead = lead(job_change, lag_months)) %>%
    filter(!is.na(cs), !is.na(jobchg_lead))
  
  cs_min   <- cs_limits[1]
  cs_max   <- cs_limits[2]
  jb_min   <- jobs_limits[1]
  jb_max   <- jobs_limits[2]
  
  cs_range <- cs_max - cs_min
  jb_range <- jb_max - jb_min
  
  df <- df %>%
    mutate(
      jobchg_scaled = cs_min + (jobchg_lead - jb_min) * (cs_range / jb_range)
    )
  
  jobs_label <- paste0("Job change (t + ", lag_months, "m)")
  
  ggplot(df, aes(x = date)) +
    geom_line(aes(y = cs, color = "Consumer Sentiment")) +
    geom_line(aes(y = jobchg_scaled, color = jobs_label)) +
    scale_y_continuous(
      name   = "Consumer Sentiment Index",
      limits = cs_limits,
      sec.axis = sec_axis(
        trans = ~ (.- cs_min) * (jb_range / cs_range) + jb_min,
        name  = paste0("Added Job (thousands) at t + ",
                       lag_months, " months")
      )
    ) +
    scale_color_manual(
      values = setNames(
        c("darkblue", "darkgreen"),
        c("Consumer Sentiment", jobs_label)
      )
    ) +
    labs(
      title = paste0("Consumer Sentiment vs Added Jobs (", lag_months, "-month lead)"),
      x     = "Date",
      color = NULL
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom"
    )
}

cs_vs_jobs(macro, lag_months =0)
cs_vs_jobs(macro, lag_months =6)
cs_vs_jobs(macro, lag_months =5)
cs_vs_jobs(macro, lag_months =7)
cs_vs_jobs(macro, lag_months =8)
```
```{r}
#| warning: false
#| message: false
# function to plot Consumer Sentiment vs Job Change with dual y-axes + smoothing
cs_vs_jobs <- function(data,
                       lag_months  = 3,
                       cs_limits   = c(40, 120),   # left y-axis
                       jobs_limits = c(-1500, 1500)) { # right y-axis (thousands)
  
  df <- data %>%
    arrange(date) %>%
    mutate(jobchg_lead = lead(job_change, lag_months)) %>%
    filter(!is.na(cs), !is.na(jobchg_lead))
  
  cs_min   <- cs_limits[1]
  cs_max   <- cs_limits[2]
  jb_min   <- jobs_limits[1]
  jb_max   <- jobs_limits[2]
  
  cs_range <- cs_max - cs_min
  jb_range <- jb_max - jb_min
  
  df <- df %>%
    mutate(
      jobchg_scaled = cs_min + (jobchg_lead - jb_min) * (cs_range / jb_range)
    )
  
  jobs_label <- paste0("Job change (t + ", lag_months, "m)")
  
  ggplot(df, aes(x = date)) +
    # raw lines (lighter)
    geom_line(aes(y = cs, color = "Consumer Sentiment"), alpha = 0.4) +
    geom_line(aes(y = jobchg_scaled, color = jobs_label), alpha = 0.4) +
    
    # smoothed lines (highlighted)
    geom_smooth(aes(y = cs, color = "Consumer Sentiment"),
                se = FALSE, span = 0.2, linewidth = 0.9) +
    geom_smooth(aes(y = jobchg_scaled, color = jobs_label),
                se = FALSE, span = 0.2, linewidth = 0.9) +
    
    scale_y_continuous(
      name   = "Consumer Sentiment Index",
      limits = cs_limits,
      sec.axis = sec_axis(
        trans = ~ (.- cs_min) * (jb_range / cs_range) + jb_min,
        name  = paste0("Added Job (thousands) at t + ",
                       lag_months, " months")
      )
    ) +
    scale_color_manual(
      values = setNames(
        c("darkblue", "darkgreen"),
        c("Consumer Sentiment", jobs_label)
      )
    ) +
    labs(
      title = paste0("Consumer Sentiment vs Added Jobs (", lag_months, "-month lead)"),
      x     = "Date",
      color = NULL
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom"
    )
}

cs_vs_jobs(macro, lag_months =0)
cs_vs_jobs(macro, lag_months =6)
cs_vs_jobs(macro, lag_months =5)
cs_vs_jobs(macro, lag_months =7)
cs_vs_jobs(macro, lag_months =8)
```


\newpage

## Visualization 2. Scatter: net expectations vs subsequent unemployment change

```{r}
#| warning: false
#| message: false
macro2 <- macro %>%
  left_join(UNEPLY_supp, by = "date") %>%
  arrange(date)

expect_vs_unemp_scatter <- function(data,
                                    horizon_months = 3,
                                    x_limits = c(-60, 60),
                                    y_limits = c(-3, 3)) {
  df <- data %>%
    mutate(
      unrate_delta = lead(unrate, horizon_months) - unrate
    ) %>%
    filter(!is.na(net_unemp_expect), !is.na(unrate_delta))
  
  # correlation and regression
  cor_val <- cor(df$net_unemp_expect, df$unrate_delta)
  
  lm_fit  <- lm(unrate_delta ~ net_unemp_expect, data = df)
  lm_sum  <- summary(lm_fit)
  
  slope   <- coef(lm_fit)[["net_unemp_expect"]]
  r2      <- lm_sum$r.squared
  
  # print a brief statistical summary in the console
  cat("\nHorizon:", horizon_months, "months\n")
  cat("Correlation (net expectations, future Δunemp):",
      round(cor_val, 3), "\n")
  cat("Slope (OLS):", round(slope, 4),
      "pp change in unemp per 1-pt net expectation\n")
  cat("R-squared:", round(r2, 3), "\n\n")
  
  # scatter plot with tuned aesthetics
  ggplot(df, aes(x = net_unemp_expect, y = unrate_delta)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey") +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey") +
    geom_point(alpha = 0.382, size = 1.618) +
    geom_smooth(method = "lm", se = FALSE, color = "firebrick", linewidth = 1) +
    scale_x_continuous(
      limits = x_limits,
      breaks = seq(x_limits[1], x_limits[2], by = 20)
    ) +
    scale_y_continuous(
      limits = y_limits,
      breaks = seq(y_limits[1], y_limits[2], by = 1)
    ) +
    labs(
      title = paste0(
        "Net unemployment expectations vs. ",
        horizon_months, "-month change in unemployment"
      ),
      subtitle = paste0(
        "Slope ≈ ", round(slope, 3),
        " pp / 1-pt net expectation,  R² ≈ ", round(r2, 2)
      ),
      x = "Net % expecting less minus more unemployment (t)",
      y = paste0("Change in unemployment over next ",
                 horizon_months, " months (percentage points)")
    ) +
    theme_minimal() +
    theme(
      plot.title.position = "plot",
      legend.position = "none"
    )
}

# explore horizons from 1 to 12 months
horizons <- 0:12

plots_set <- vector("list", length(horizons))
names(plots_set) <- paste0("h_", horizons, "m")

for (i in seq_along(horizons)) {
  plots_set[[i]] <- expect_vs_unemp_scatter(
    macro2,horizons[i])
}

plots_set
```

## Visualization 3. Time series: net expectations vs actual unemployment (dual axis)

```{r}
#| warning: false
#| message: false
expect_vs_unemp_dual <- function(data_macro2,
                                 cs_limits    = c(-60, 60),# for net expectations
                                 unemp_limits = c(2, 12)) {# for unemployment
  
  df <- data_macro2 %>%
    filter(!is.na(net_unemp_expect), !is.na(unrate)) %>%
    arrange(date)
  
  ex_min <- cs_limits[1]
  ex_max <- cs_limits[2]
  ur_min <- unemp_limits[1]
  ur_max <- unemp_limits[2]
  
  ex_rng <- ex_max - ex_min
  ur_rng <- ur_max - ur_min
  
  df <- df %>%
    mutate(
      unrate_scaled = ex_min + (unrate - ur_min) * (ex_rng / ur_rng)
    )
  
  ggplot(df, aes(x = date)) +
    # raw lines (lighter)
    geom_line(aes(y = net_unemp_expect,
                  color = "Net expectations (Less-More)"),
              alpha = 0.4) +
    geom_line(aes(y = unrate_scaled,
                  color = "Unemployment"),
              alpha = 0.4) +
    
    # smoothed lines (highlighted)
    geom_smooth(aes(y = net_unemp_expect,
                    color = "Net expectations (Less-More)"),
                se = FALSE, span = 0.2, linewidth = 0.9) +
    geom_smooth(aes(y = unrate_scaled,
                    color = "Unemployment"),
                se = FALSE, span = 0.2, linewidth = 0.9) +
    
    scale_y_continuous(
      name   = "Net (Less-More) unemployment expectations",
      limits = cs_limits,
      sec.axis = sec_axis(
        trans = ~ (.- ex_min) * (ur_rng / ex_rng) + ur_min,
        name  = "Unemployment rate (%)"
      )
    ) +
    scale_color_manual(
      values = setNames(
        c("darkgreen", "firebrick"),
        c("Net expectations (Less-More)", "Unemployment")
      )
    ) +
    labs(
      title = "Net Unemployment Expectations vs Actual Unemployment",
      x     = "Date",
      color = NULL
    ) +
    theme_minimal() +
    theme(legend.position = "bottom")
}


expect_vs_unemp_dual(macro2)
```
```{r}
#| warning: false
#| message: false
expect_vs_unemp_dual <- function(data_macro2,
                                 lag_months   = 3,
                                 cs_limits    = c(-60, 60),  # for net expectations
                                 unemp_limits = c(2, 12)) {  # for unemployment
  
  df <- data_macro2 %>%
    arrange(date) %>%
    mutate(unrate_lead = lead(unrate, lag_months)) %>%
    filter(!is.na(net_unemp_expect), !is.na(unrate_lead))
  
  ex_min <- cs_limits[1]
  ex_max <- cs_limits[2]
  ur_min <- unemp_limits[1]
  ur_max <- unemp_limits[2]
  
  ex_rng <- ex_max - ex_min
  ur_rng <- ur_max - ur_min
  
  df <- df %>%
    mutate(
      unrate_scaled = ex_min + (unrate_lead - ur_min) * (ex_rng / ur_rng)
    )
  
  unemp_label <- paste0("Unemployment (t + ", lag_months, "m)")
  
  ggplot(df, aes(x = date)) +
    # raw lines (lighter)
    geom_line(aes(y = net_unemp_expect,
                  color = "Net expectations (Less-More)"),
              alpha = 0.4) +
    geom_line(aes(y = unrate_scaled,
                  color = unemp_label),
              alpha = 0.4) +
    
    # smoothed lines (highlighted)
    geom_smooth(aes(y = net_unemp_expect,
                    color = "Net expectations (Less-More)"),
                se = FALSE, span = 0.2, linewidth = 0.9) +
    geom_smooth(aes(y = unrate_scaled,
                    color = unemp_label),
                se = FALSE, span = 0.2, linewidth = 0.9) +
    
    scale_y_continuous(
      name   = "Net (Less-More) unemployment expectations",
      limits = cs_limits,
      sec.axis = sec_axis(
        trans = ~ (.- ex_min) * (ur_rng / ex_rng) + ur_min,
        name  = paste0("Unemployment rate (%) at t + ",
                       lag_months, " months")
      )
    ) +
    scale_color_manual(
      values = setNames(
        c("darkgreen", "firebrick"),
        c("Net expectations (Less-More)", unemp_label)
      )
    ) +
    labs(
      title = paste0("Net Unemployment Expectations vs Future Unemployment (",
                     lag_months, "-month lead)"),
      x     = "Date",
      color = NULL
    ) +
    theme_minimal() +
    theme(legend.position = "bottom")
}

expect_vs_unemp_dual(macro2, lag_months = 3)
expect_vs_unemp_dual(macro2, lag_months = 6)
expect_vs_unemp_dual(macro2, lag_months = 5)
expect_vs_unemp_dual(macro2, lag_months = 7)
expect_vs_unemp_dual(macro2, lag_months = 8)
expect_vs_unemp_dual(macro2, lag_months = 9)
expect_vs_unemp_dual(macro2, lag_months = 10)
expect_vs_unemp_dual(macro2, lag_months = 11)
expect_vs_unemp_dual(macro2, lag_months = 12)
```

\newpage

# Indications

**SOME REAL WORLD INDICATIONS.**

\newpage

# Conclusion & Outlook

| 

# Limitation

|

\newpage

# References

https://data.sca.isr.umich.edu/data-archive/mine.php

https://www.bls.gov/cps

https://www.bls.gov/ces
